Site-Heterogeneous Character Change Models for Morphology

### 2018-10-30 Papers/DraftOutline_SSCCM.tex
adding some bibliography stuff

\documentclass[]{article}
\usepackage{amsmath}

@@

\subsubsection{Discretized beta model}
We model site-specific state frequencies by assuming the state frequency $\pi_i$ at site $i$ is drawn from a Beta prior distribution with hyperparameters $\alpha,\beta$.
Then, the likelihood of the data $D$ is computed by marginalizing over the value for $\pi_i$ at each site.

\begin{equation}\label{continuous-likelihood}
f(D\mid \alpha,\beta) = \prod_{i=1}^n\int_0^1 f(D_i \mid \pi_i)f(\pi_i \mid \alpha,\beta)d\pi_i
\end{equation}

To simplify the computation, we approximate this integrated likelihood by assuming $\pi_i$ is drawn from a mixture distribution, where the mixture categories are defined deterministically by discretizing a Beta distribution into $k$ bins whose boundaries are defined using $k-1$ quantiles.
The the value of mixture category $j$ is indicated by $\beta_j$ and is computed as the interquantile median of the $j$-th bin.
In other words, $\beta_j = I^{-1}_{(2j-1)/2k}(\alpha,\beta)$, where $I_x(\alpha,\beta)$ is the regularized incomplete Beta function.
The mixing proportion is $1/k$ for each category.
Then the likelihood is computed by summing over the $k$ discrete mixture categories for $\pi_i$ at each site.

\begin{equation}\label{discretized-likelihood}
f(D\mid \alpha,\beta) = \prod_{i=1}^n\frac{1}{k}\sum_{j=1}^k f(D_i \mid \pi_i = \beta_j)
\end{equation}

@@

\subsubsection{Site-heterogeneous discrete multistate model}

@@

@@

>>>>

\documentclass[]{article}
\usepackage{amsmath} 
\usepackage{natbib}

@@

We model morphological character evolution using the Mk model \citep{lewis2001likelihood}, and additionally account for site-specific heterogeneity in substitution rates among character states by allowing the equilibrium state frequencies $\pi$ to vary among sites, such that each site $i$ has its own state frequency vector $\pi_i$.

\subsubsection{Discretized beta model}
We model site-specific binary state frequencies by assuming the state frequencies $\pi_i$ are drawn from a 2-dimensional Dirichlet (Beta) prior distribution with hyperparameters $\alpha,\beta$.
Then, the likelihood of the data $D$ is computed by marginalizing over the value for $\pi_i$ at each site.

\begin{equation} \label{continuous-likelihood}
f(D\mid \alpha,\beta) = \prod_{i=1}^n\int_0^1 f(D_i \mid \pi_i)f(\pi_i \mid \alpha,\beta)d\pi_i
\end{equation}

To simplify the computation, we approximate this integrated likelihood by assuming $\pi_i$ is drawn from a mixture distribution, where the mixture categories are defined deterministically by discretizing a Beta distribution into $k$ bins whose boundaries are defined using $k-1$ quantiles.
The the value of mixture category $j$ is indicated by $\phi_j$ and is computed as the interquantile median of the $j$-th bin.
In other words, $\phi_j = I^{-1}_{(2j-1)/2k}(\alpha,\beta)$, where $I_x(\alpha,\beta)$ is the regularized incomplete Beta function.
The mixing proportion is $1/k$ for each category.
Then the likelihood is computed by summing over the $k$ discrete mixture categories for $\pi_i$ at each site.

\begin{equation} \label{discretized-likelihood}
f(D\mid \alpha,\beta) = \prod_{i=1}^n\frac{1}{k}\sum_{j=1}^k f(D_i \mid \pi_i = \phi_j)
\end{equation}

@@

\subsubsection{Site-heterogeneous multistate model}

@@

\subsubsection{Model comparison}

The appropriate number of categories $k$ is selected by comparing estimates of the marginal likelihood under different values for $k$.
Marginal likelihoods were estimated using the steppingstone sampling method \citep{xie2010improving} implemented in RevBayes.

@@

\bibliographystyle{plain}
\bibliography{bibliography}
